Basic concepts and tools for the Toki Pona minimalist language: Wordnet synsets; analysis, synthesis and syntax highlighting of texts

### 2017-12-15 article/pos.tex
terminada revisao

\begin{table*}[h!]
\begin{center}
\caption{"POS tags incident and chosen.
            The official dictionary often relates tokens
            to more than one POS tag.
            For the text highlighting Plugin, for example,
            a token has to have an established tag to have
            a defined color.
            On the Chosen column, the tokens were regarded only once
            by choosing the first classification in the dictionary in ['PRE', 'VERB', 'PREPOSITION', 'PARTICLE', 'ADJECTIVE', 'NOUN', 'NUMBER'].}
\begin{tabular}{| l | c | c |}\hline
POS & All  & Chosen \\\hline
NOUN & 58  & 49 \\\hline
ADJECTIVE & 40  & 37 \\\hline
VERB & 15  & 15 \\\hline
PARTICLE & 12  & 12 \\\hline
PREPOSITION & 5  & 5 \\\hline
NUMBER & 4  & 1 \\\hline
PRE & 0  & 0 \\\hline
total & 134  & 120 \\\hline
\end{tabular}\end{center}
\end{table*}

>>>>

\begin{table*}[h!]
\begin{center}
\caption{POS tags incident and chosen.
            The official dictionary often relates tokens
            to more than one POS tag.
            For the current version of the text highlighting plugin described in
            Section~\ref{shigh}, for example,
            a token has to have an established tag to have
            a defined color.
            On the 'Chosen' column, the tokens were regarded only once
            by choosing the class in the dictionary respecting the
            precedence order: PRE, VERB, PREPOSITION, PARTICLE, ADJECTIVE, NOUN, NUMBER.}\label{foobar}
\begin{tabular}{| l | c | c |}\hline
POS & All  & Chosen \\\hline
NOUN & 58  & 49 \\\hline
ADJECTIVE & 40  & 34 \\\hline
VERB & 15  & 13 \\\hline
PARTICLE & 12  & 12 \\\hline
PRE & 6  & 6 \\\hline
PREPOSITION & 5  & 5 \\\hline
NUMBER & 4  & 1 \\\hline
total & 140  & 120 \\\hline
\end{tabular}\end{center}
\end{table*}

### 2017-12-15 article/vowels.tex
terminada revisao

\begin{table*}[h!]
\begin{center}
\caption{"Frequency of vowels in Toki Pona.}
\begin{tabular}{| l | c |}\hline
Vowel & freq \\\hline
a & 33.19 \\\hline
e & 17.45 \\\hline
i & 23.40 \\\hline
o & 15.32 \\\hline
u & 10.64 \\\hline
\end{tabular}\end{center}
\end{table*}

>>>>

\begin{table*}[h]
\scriptsize
\begin{center}
\caption{Frequency of letters in Toki Pona.
            freq, freq\_I, freq\_L and freq\_M are
            the frequencies of the letters in any, initial, last and middle
            positions.
            The columns 'v' and 'c' that follow them are frequencies
            considering only vowels and consonants.
            The most frequent vowel is 'a' in any position,
            although it is more salient among words starting with a vowel
            and among the last letter of the words.
            For starting, ending and middle positions, the second most frequent
            vowel varies.
            Among the consonants, 'n' is the most frequent because it is
            the only consonant allowed in the last position and because
            almost 20\% of the words end with 'n'.
            On the initial position, 's' is the most frequent consonant,
            while in middle position 'l' is the most frequent consonant.
            Many other conclusions can be drawn from this table and are
            useful e.g. for exploring sonorities in poems.}\label{freqLet}
\hspace*{-1.3cm}
\begin{tabular}{  l | c   c   c | c   c   c | c   c   c | c   c   c  }
letter & freq  & v  & c  & freq\_I  & v  & c  & freq\_L  & v  & c  & freq\_M  & v  & c \\\hline
a & 16.35  & 33.19  & -  & 8.06  & 40.00  & -  & 29.03  & 35.64  & -  & 14.22  & 29.46  & - \\
e & 8.60  & 17.45  & -  & 2.42  & 12.00  & -  & 11.29  & 13.86  & -  & 10.78  & 22.32  & - \\
i & 11.53  & 23.40  & -  & 3.23  & 16.00  & -  & 20.97  & 25.74  & -  & 10.78  & 22.32  & - \\
o & 7.55  & 15.32  & -  & 4.03  & 20.00  & -  & 14.52  & 17.82  & -  & 6.03  & 12.50  & - \\
u & 5.24  & 10.64  & -  & 2.42  & 12.00  & -  & 5.65  & 6.93  & -  & 6.47  & 13.39  & - \\\hline
j & 2.10  & -  & 4.13  & 3.23  & -  & 4.04  & 0.00  & -  & 0.00  & 2.59  & -  & 5.00 \\
k & 6.29  & -  & 12.40  & 11.29  & -  & 14.14  & 0.00  & -  & 0.00  & 6.90  & -  & 13.33 \\
l & 9.22  & -  & 18.18  & 12.10  & -  & 15.15  & 0.00  & -  & 0.00  & 12.50  & -  & 24.17 \\\hline
m & 4.61  & -  & 9.09  & 10.48  & -  & 13.13  & 0.00  & -  & 0.00  & 3.88  & -  & 7.50 \\
n & 10.48  & -  & 20.66  & 6.45  & -  & 8.08  & 18.55  & -  & 100.00  & 8.19  & -  & 15.83 \\
p & 5.66  & -  & 11.16  & 11.29  & -  & 14.14  & 0.00  & -  & 0.00  & 5.60  & -  & 10.83 \\\hline
s & 6.29  & -  & 12.40  & 13.71  & -  & 17.17  & 0.00  & -  & 0.00  & 5.60  & -  & 10.83 \\
t & 3.14  & -  & 6.20  & 6.45  & -  & 8.08  & 0.00  & -  & 0.00  & 3.02  & -  & 5.83 \\
w & 2.94  & -  & 5.79  & 4.84  & -  & 6.06  & 0.00  & -  & 0.00  & 3.45  & -  & 6.67 \\
\end{tabular}\end{center}
\end{table*}
